Concurrent Programming in Harmony

### 2021-08-09 book/paper.tex
Latest
Improved Bosco code

\begin{figure}
\begin{code}
\harmonysource{chain}
\end{code}
\caption{\harmonylink{code/chain.hny} Checking the correctness of Chain Replication.}
\label{fig:chain}
\end{figure}

@@

\autoref{fig:bosco} presents our algorithm.  Besides the network variable, it uses a
shared list of proposals and a shared set of decisions.
In this particular algorithm, all messages are broadcast to all processors, so they do
not require a destination address.
The invariants we want to maintain
on those variables is that $\textit{decisions} \subseteq \textit{proposals}$
and $|\textit{decisions}| \leq 1$.  Since the initial value of \textit{decisions} is
$\emptyset$, the invariants clearly hold initially.

@@

In Line~21, a processor waits for $\texttt{N} - \texttt{F}$ messages using
the Harmony \textbf{select} statement.
Since Harmony has to check all possible executions of the protocol, the
\texttt{receive}(\textit{round}, $k$) method returns all \emph{subbags} of messages
for the given round that have size $k = \texttt{N} - \texttt{F}$.  The method uses a
dictionary comprehension to filter out all messages for the given \textit{round}
and then uses the \texttt{bag.combinations} method to find all combinations of size $k$.
The \textbf{select} statement waits until there is at least one such combination and
then chooses an element, which is bound to the \textit{quorum} variable.
The body of the \textbf{select} statement is then executed atomically.
This is usually how distributed algorithms are modeled, because they can only interact
through the network.  There is no need to interleave the different processes other
than when messages are delivered.  By executing the body atomically, a lot of
unnecessary interleavings are avoided and this reduces the state space that must
be explored by the model checker significantly.

@@

While one can run this code within little time for $\mathtt{F} = 1$, for
$\mathtt{F} = 2$ the state space to explore is already quite large.
One way to reduce the state space to explore is the following realization:
each processor only considers messages for the round that it is in.
If a message is for an old round, the processor will ignore it;
if a message is for a future round, the processor will buffer it.
So one can simplify the model and have each processor wait
for \emph{all} \texttt{N} messages in a round
instead of $\mathtt{N} - \mathtt{F}$.
It would still have to choose to consider just $\mathtt{N} - \mathtt{F}$
out of those \texttt{N} messages, but executions in which some processors
are left behind in all rounds are no longer considered.
It still includes executions where some subset of $\mathtt{N} - \mathtt{F}$
processors only choose each other messages and essentially ignore the
messages of the remaining \texttt{F} processors, so the resulting model
is just as good.

\autoref{fig:bosco2} shows the code for this model.  Running this with
$\mathtt{F} = 2$ does not take very long and this approach is a good
blueprint for testing other round-based protocols (of which there are
many).

>>>>

\begin{figure}
\begin{code}
\harmonysource{chain}
\end{code}
\caption{\harmonylink{code/chain.hny} Checking correctness of Chain Replication.}
\label{fig:chain}
\end{figure}

@@

\autoref{fig:bosco} presents our algorithm.  Besides the network variable, it uses a
shared list of proposals and a shared set of decisions.
To reduce the state space to explore, not all permutations of zeroes and ones are
explored.  With 5 processors ($\mathtt{F} = 2)$, say, we only explore the cases where
no processors propose 1, where exactly one processors proposes 1, and where 2 processors
proposes 1.
In this particular algorithm, all messages are broadcast to all processors, so they do
not require a destination address.
The invariants we want to maintain
on those variables is that $\textit{decisions} \subseteq \textit{proposals}$
and $|\textit{decisions}| \leq 1$.  Since the initial value of \textit{decisions} is
$\emptyset$, the invariants clearly hold initially.

@@

In Line~22, a processor waits for $\texttt{N} - \texttt{F}$ messages using
the Harmony \textbf{select} statement.
Since Harmony has to check all possible executions of the protocol, the
\texttt{receive}(\textit{round}, $k$) method returns all \emph{subbags} of messages
for the given round that have size $k = \texttt{N} - \texttt{F}$.  The method uses a
dictionary comprehension to filter out all messages for the given \textit{round}
and then uses the \texttt{bag.combinations} method to find all combinations of size $k$.
The \textbf{select} statement waits until there is at least one such combination and
then chooses an element, which is bound to the \textit{quorum} variable.
The body of the \textbf{select} statement is then executed atomically.
This is usually how distributed algorithms are modeled, because they can only interact
through the network.  There is no need to interleave the different processes other
than when messages are delivered.  By executing the body atomically, a lot of
unnecessary interleavings are avoided and this reduces the state space that must
be explored by the model checker significantly.

@@

While one can run this code within little time for $\mathtt{F} = 2$, for
$\mathtt{F} = 3$ the state space to explore is already quite large.
One way to reduce the state space to explore is the following realization:
each processor only considers messages for the round that it is in.
If a message is for an old round, the processor will ignore it;
if a message is for a future round, the processor will buffer it.
So one can simplify the model and have each processor wait
for \emph{all} \texttt{N} messages in a round
instead of $\mathtt{N} - \mathtt{F}$.
It would still have to choose to consider just $\mathtt{N} - \mathtt{F}$
out of those \texttt{N} messages, but executions in which some processors
are left behind in all rounds are no longer considered.
It still includes executions where some subset of $\mathtt{N} - \mathtt{F}$
processors only choose each other messages and essentially ignore the
messages of the remaining \texttt{F} processors, so the resulting model
is just as good.

\autoref{fig:bosco2} shows the code for this model.  Running this with
$\mathtt{F} = 3$ does not take very long and this approach is a good
blueprint for testing other round-based protocols (of which there are
many).
